---
title: "INPC Social Network Analysis"
author: "Daniel P. Hall Riggins, MD, MPH"
date: today
format: html
toc: true
code-fold: true
embed-resources: true
df-print: paged
execute: 
  cache: true
---

```{r}
#| output: false

# Libraries
library(targets)
library(tidyverse)
library(janitor)
library(tidygraph)
library(ggraph)
library(networkD3)

# Custom Functions
sapply(
    paste0("R/", list.files("R/")),
    source
)

# Data
tar_load(coauthor_network)

coauthor_nodes <-
    coauthor_network |> 
    activate(nodes) |> 
    as_tibble() |>
    mutate(
        institution_display_name = case_match(
            institution_display_name,
            "Indiana University – Purdue University Indianapolis" ~ "IUPUI",
            "Purdue University West Lafayette" ~ "Purdue",
            "United States Department of Veterans Affairs" ~ "Veterans Affairs",
            .default = institution_display_name
        )
    ) |> 
    select(
        Author = au_display_name, 
        Institution = institution_display_name,
        Degree = degree, 
        Eigencentrality = eigen, 
        Betweeness = betweeness,
        Cluster = louvain_group
    )
```

For this analysis, I have converted the dataset into a network, where each node of the network represents an author and each connection between nodes represents a paper (or other scholarly work) on which two authors collaborated.

For example, here is the local network centered around Brian Dixon and all his coauthors. You can use your mouse to pan around and zoom in/out:


```{r}
brian_network <-
    coauthor_network |> 
    activate(nodes) |>
    convert(
        to_local_neighborhood,
        node = 19
    ) 

brian_nodes <-
    brian_network |>  
    as_tibble() |> 
    as.data.frame()

brian_edges <- 
    brian_network |> 
    activate(edges) |> 
    as_tibble() |> 
    mutate(
        from = from - 1,
        to = to - 1
    ) |> 
    as.data.frame()

forceNetwork(
    Links = brian_edges,
    Nodes = brian_nodes,
    Source = "from",
    Target = "to",
    NodeID = "au_display_name",
    Group = "institution_display_name",
    zoom = TRUE,
    opacity = 0.9,
    opacityNoHover = 0.5,
    # legend = TRUE
    # charge = -30
)
```


# Degree

This metric measures the number of links connected to each node. In this context it reflects the number of coauthors each person has shared a paper with.

```{r}
coauthor_nodes |> 
    slice_max(order_by = Degree, n = 10)
```

# Eigencentrality

This metric is like Degree, but also takes into account the degree of connectedness throughout the broader network, so it can give a more gestalt view of who is most central to the network.

```{r}
coauthor_nodes |> 
    relocate(Eigencentrality, .before = Degree) |> 
    slice_max(order_by = Eigencentrality, n = 10)
```

# Betweeness

This metric measures how much nodes act as bridges between different clusters of a network. In this context it reflects which authors act as connecters between different clusters of collaborators.

```{r}
coauthor_nodes |> 
    relocate(Betweeness, .before = Degree) |> 
    slice_max(order_by = Betweeness, n = 10)
```


# Clusters

The Louvain algorithm is a way of automatically detecting clusters within a network. Unlike many other cluster-detection algorithms, it has the advantage of not requiring you to pre-specify how many clusters you think there should be. See below which cluster each author has been assigned to:

```{r}
coauthor_nodes |> 
    relocate(Cluster, .before = Degree) |> 
    arrange(Cluster, Author)
```

And here those clusters are visualized:

```{r}
cluster_nodes <-
    coauthor_network |> 
    activate(nodes) |>
    as_tibble() |> 
    as.data.frame()

cluster_edges <- 
    coauthor_network |> 
    activate(edges) |> 
    as_tibble() |> 
    mutate(
        from = from - 1,
        to = to - 1
    ) |> 
    as.data.frame()

forceNetwork(
    Links = cluster_edges,
    Nodes = cluster_nodes,
    Source = "from",
    Target = "to",
    NodeID = "au_display_name",
    Group = "louvain_group",
    zoom = TRUE,
    legend = TRUE,
    opacity = 0.9,
    opacityNoHover = 0.75,
    charge = -30
)
```

