---
title: "Initializing the Dolt Database of Regenstrief Works pulled from OpenAlex"
author: "Daniel P. Hall Riggins, MD, MPH"
format: html
---

Start by loading dependencies:

```{r}
#library(doltr)
library(openalexR)
library(tidyverse)
```

I started a [database on DoltHub](https://www.dolthub.com/repositories/danriggins/regenstrief_from_openalex) and here I pull it to my local computer as one would with a git repository.

```{r}
#dolt_clone("danriggins/regenstrief_from_openalex")
Sys.setenv(DOLT_DIR="regenstrief_from_openalex")
```

Next I queried the OpenAlex API for all works affiliated with Regenstrief Institute, which were returned as a nested R dataframe.

```{r}
regenstrief_works <-
    oa_fetch(
        entity = "works",
        institutions.ror = "https://ror.org/05f2ywb48",
        verbose = TRUE
    )
```

You can't write a nested dataframe to a SQL database directly (as far as I know), so I flattened out the object into two-dimensional tables, starting with the core table, which consists of all "works" (which are mostly peer-reviewed journal articles).

```{r}
regenstrief_works_2d <-
    regenstrief_works |> 
    select(
        !where(is.list)
    ) |> 
    mutate(
        publication_date = ymd(publication_date)
    ) |> 
    rename(work_id = id)
```

And here is the code for writing the table to the Dolt database (commented out so it doesn't run a second time). 

```{r}
# dbWriteTable(dolt(), "works", regenstrief_works_2d)
# dolt_status()
# dolt_add("works")
# dolt_commit("Adding core works table")
# dolt_push()
```

You can see all the additional nested tables using this query:

```{r}
regenstrief_works |> 
    select(
        where(is.list)
    ) |> 
    glimpse()
```

Next, we repeat the process for all nested tables:

```{r}
funders <-
    regenstrief_works |> 
    select(
        work_id = id,
        grants
    ) |> 
    filter(!is.na(grants)) |>
    unnest(grants) |> 
    mutate(
        col_names = 
            rep_len(
                c("funder_id", "funder_display_name", "award_id"), 
                3198
            )
    ) |>
    pivot_wider(
        names_from = col_names, 
        values_from = grants
    ) |>
    unnest(funder_id, funder_display_name, award_id) 
    
# dbWriteTable(dolt(), "funders", funders)
# dolt_status()
# dolt_add("funders")
# dolt_commit("Adding funders table")
# dolt_push

work_id_lookup <-
    regenstrief_works |> 
    select(
        work_id = id,
        ids
    ) |> 
    unnest(ids) |> 
    mutate(
        type = 
            case_when(
                str_detect(ids, "^https://open") ~ "openalex",
                str_detect(ids, "^https://doi") ~ "doi",
                str_detect(ids, "^https://www.ncbi") ~ "pmc_id",
                str_detect(ids, "^https://pubmed") ~ "pubmed_id",
                str_detect(ids, "^[0-9]") ~ "mag_id"
            )
    ) |> 
    filter(type != "openalex") |> 
    filter(type != "doi") |> 
    pivot_wider(names_from = type, values_from = ids)

regenstrief_works_2d_with_ids <-
    full_join(work_id_lookup, regenstrief_works_2d, by = "work_id")

# dbWriteTable(dolt(), "works", regenstrief_works_2d_with_ids, overwrite = TRUE)
# dolt_status()
# dolt_add("works")
# dolt_commit(message = "Adding additional ids to works")
# dolt_push()

authorships <-
    regenstrief_works |> 
    select(
        work_id = id,
        author
    ) |> 
    unnest(author)

# dbWriteTable(dolt(), "authorships", authorships)
# dolt_add("authorships")
# dolt_commit(message = "Adding authorships")
# dolt_pull()
# dolt_push()
# dolt_log()
# dolt_status()

referenced_works <-
    regenstrief_works |> 
    select(
        from_work_id = id,
        referenced_works
    ) |> 
    unnest(referenced_works) |> 
    rename(to_work_id = referenced_works)

# dbWriteTable(dolt(), "referenced_works", referenced_works, overwrite = TRUE)
# dolt_status()
# dolt_add("referenced_works")
# dolt_commit(message = "Adding referenced works")
# dolt_push()

library(openalexR)

oa_fetch(
            entity = "works",
            from_publication_date = "2013-07-01", # Since cut off date for prior review
            to_publication_date = "2023-03-31", # Through first quarter of 2023
            search = "\"indiana network for patient care\""
        )

related_works <-
    regenstrief_works |> 
    select(
        from_work_id = id,
        related_works
    ) |> 
    unnest() |> 
    rename(to_work_id = related_works)

# dbWriteTable(dolt(), "related_works", related_works, overwrite = TRUE)
# dolt_status()
# dolt_add("related_works")
# dolt_commit(message = "Adding related works")
# dolt_push()

work_concepts <-
    regenstrief_works |> 
    select(
        work_id = id,
        concepts
    ) |> 
    unnest(concepts) |> 
    rename(concept_id = id)

dbWriteTable(dolt(), "work_concepts", work_concepts, overwrite = TRUE)
dolt_status()
dolt_add("work_concepts")
dolt_commit(message = "Adding work concepts")
dolt_push()
```